---
title: "Meeting Client Needs while Maintaining Spatial Integrity:"
subtitle: "A Spatial Project on Digitalization in Farming"
author: "Mia Thuge Kuntz"
date: "12-May-2023 updated `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(knitr)

## global options
options(max.print="75")
opts_chunk$set(cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r remove, include=FALSE}
rm(list=ls())
```

## Preparing Data
In order to visualise the data in a meaningful way I first need to prepare it. First step is reading in the necessary libraries and then read in the data and inspect it.

```{r libraries}
# loading libraries
library(googlesheets4)
library(leaflet)
library(sf)
library(RColorBrewer)
library(dplyr)
library(htmltools)
library(tidyverse)
library(ggplot2)
library(ggnewscale)
```

```{r read data}
# getting authentication
gs4_deauth()
# reading file containing articles' information
articles <- read_sheet("https://docs.google.com/spreadsheets/d/1NFlbYvgNJCsr0uW5uyhvtZPhoj4UlpB0D9DB7rlhu6g/edit?usp=sharing", range = "Ark1")
articles

# reading file containing articles' information (offline access)
# library(readxl)
# articles <- read_excel("data/articles.xlsx")

# reading GeoJSON file containing country boundaries
countries <- st_read("data/countries.geojson")
countries
```

The `articles` data is read in as a data frame, and at first inspection seems to have everything it should have included. The `countries` file is read in as a simple feature object with 255 features and 2 fields. Its a multipolygon, which means that the countries geometry is a collection of multiple polygons. The dimension of the data is a two dimensional coordinate system "XY", where the bounding box values show that it covers nearly the whole world. 
The coordinate reference system (CRS) of the data is WGS84, which is typical when working with global spatial data.  

Next part of the code focuses on preparing the data to be visualised. I first split those rows in the "Country" column which has several countries in it so to be able to credit all countries responsible for that particular article when they are visualised in the maps. I thereafter unnest those arrays into their own set of rows. Lastly, I merge the two data set into one `merged_data` data frame by the two columns where they share content.     

```{r prep data}
# splitting multiple countries in "Country" column into separate rows
articles$Country <- strsplit(articles$Country, ", ")

# unnesting to convert array into set of rows 
articles <- unnest(articles, Country)

# merging articles and countries based on country column
merged_data <- merge(articles, countries, by.x = "Country", by.y = "ADMIN", all.x = TRUE)
head(merged_data, n=3)
```

When inspecting the `merged_data` data frame it appears as if all the columns from the two data files have merged successfully. This data frame only contains the countries and ISO codes of the countries which appears in the `articles` file.

To get a quick impression of the distribution of articles across countries I wish to add a count bar for the top five countries appearing in the `articles` data. To do this I first create a separate table for the `Country` column in the `merged_data` data frame. I then convert this to a data frame of its own, so that I am able to use these counts for later use in the maps. Lastly, I arrange the `counts_df` in descending order according to the `Count` column and then assign the top five countries to its own object. This object is presented as a bar chart for a quick view of which countries has the highest count.

```{r frequency count}
# calculating frequency count of each country
country_counts <- table(merged_data$Country)

# converting frequency counts to data frame
counts_df <- data.frame(Country = names(country_counts), Count = as.numeric(country_counts))
counts_df

# merging country counts with countries GeoJSON data
merged_geojson <- merge(countries, counts_df, by.x = "ADMIN", by.y = "Country", all.x = TRUE)
merged_geojson

# getting top five countries with highest number of studies
top_countries <- counts_df %>%
  # arranging in descending order
  arrange(desc(Count)) %>%
  # selecting top five countries
  head(5)

# converting 'Country' column to factor with ordered levels based on 'Count'
top_countries$Country <- factor(top_countries$Country, levels = top_countries$Country[order(top_countries$Count)])

# plotting bar chart of top five countries count
ggplot(top_countries, aes(x = Country, y = Count)) +
  # adding bar chart
  geom_bar(stat = "identity", fill = "darkgreen") +
  # adding labels
  labs(title = "Top Five Countries Count", x = "Country", y = "Count") +
  # setting theme
  theme_bw() +
  # rotating x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # setting y-axis breaks and labels
  scale_y_continuous(breaks = seq(0, max(top_countries$Count), by = 2), labels = function(x) as.character(as.integer(x)))
```

This shows, that the country with the highest count is Germany at 12, while the "lowest" in the top five is Denmark with 3. 

## Maps
The next part of the code focuses on creating the maps. I have chosen to create two different maps, one choropleth map and one centroid map. For further explanation on the maps, please see the paper for this project.

### Choropleth Map 1

```{r choropleth prep 1}
# defining color palette
# exluding first three colors as they are too light
color_palette <- brewer.pal(9, "Greens")[-(1:3)]

# creating leaflet map with merged data
choropleth_map1 <- leaflet(data = merged_geojson) %>%
  # setting view to center of the world
  setView(lng = 0, lat = 0, zoom = 2) %>%
  # adding OpenStreetMap tiles
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  # adding polygons with color based on count of articles in each country
  addPolygons(fillColor = ~colorNumeric(color_palette, domain = Count)(Count),
              # setting weight of borders to 1
              weight = 1,
              # setting opacity of borders to 1
              opacity = 1,
              # setting color of borders to white
              color = "white",
              # setting opacity of fill color of polygons to 0.7
              fillOpacity = 0.7,
              # setting highlight options
              highlight = highlightOptions(
                # setting weight of borders to 2
                weight = 2,
                # setting color of borders to white
                color = "white",
                # setting opacity of fill color of polygons to 0.9
                fillOpacity = 0.9
              ),
              # adding labels to polygons with country name and count of articles
              label = ~paste(ADMIN, ": ", Count))
```

```{r choropleth map 1}
# displaying first choropleth map
choropleth_map1
```

### Choropleth Map 2

```{r choropleth prep 2}
# creating leaflet map with merged data
choropleth_map2 <- leaflet(data = merged_geojson) %>%
  # setting view to center of the world
  setView(lng = 0, lat = 0, zoom = 2) %>%
  # adding OpenStreetMap tiles
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  # adding polygons with color based on count of articles in each country
  addPolygons(fillColor = ~colorNumeric(color_palette, domain = Count)(Count),
              # setting weight of borders to 1
              weight = 1,
              # setting opacity of borders to 1
              opacity = 1,
              # setting color of borders to white
              color = "white",
              # setting opacity of fill color of polygons to 0.7
              fillOpacity = 0.7,
              # setting highlight options
              highlight = highlightOptions(
                # setting weight of borders to 2
                weight = 2,
                # setting color of borders to white
                color = "white",
                # setting opacity of fill color of polygons to 0.9
                fillOpacity = 0.9
              ),
              # adding labels to polygons with country name and count of articles
              label = ~paste(ADMIN, ": ", Count)) %>%
  # adding legend to bottom right corner with color palette and count values as labels 
  addLegend("bottomright", pal = colorNumeric(color_palette, domain = unique(counts_df$Count)), values = unique(counts_df$Count), title = "Count") %>%
  # adding title to top right corner with information about the map and a note about the hover function 
  addControl(html = as.character(tags$div(style = "text-align: center; background-color: white; padding: 10px; font-family: Arial, sans-serif; font-size: 16px; font-weight: bold;", 
                                          HTML(paste0("World map of studies in Smart Farming<br>",
                                                      "<span style='font-size: 12px;'>Hover mouse over country to view its count</span>")))),
              position = "topright") %>%
  # adding table to bottom left corner with top five countries and their count
  addControl(html = as.character(tags$div(HTML(paste0("<h4>Top 5 Countries</h4>",
                                                      "<table>",
                                                      "<thead><tr><th>Country</th><th>Count</th></tr></thead>",
                                                      "<tbody>",
                                                      paste0("<tr><td>", top_countries$Country, "</td><td>", top_countries$Count, "</td></tr>", collapse = "\n"),
                                                      "</tbody>",
                                                      "</table>")))),
              position = "bottomleft")
```

```{r choropleth map 2}
# displaying second choropleth map
choropleth_map2
```

### Centroid Map

```{r centroid data}
# repairing invalid geometries in merged data 
merged_geojson <- st_make_valid(merged_geojson)

# simplifying country geometries to reduce file size and increase speed of plotting 
simplified_geojson <- st_simplify(merged_geojson, preserveTopology = TRUE, dTolerance = 0.01)

# creating centroids for each country 
centroids <- st_centroid(simplified_geojson)

# extracting coordinates from centroids
centroids <- st_coordinates(centroids) 

# converting to data frame 
centroids_df <- as.data.frame(centroids)

# renaming the columns 
colnames(centroids_df) <- c("x", "y")

# combining with the count data 
centroids_df <- cbind(centroids_df, Count = simplified_geojson$Count)
```

```{r centroid map prep}
# creating centroid map 
centroid_map <- ggplot() +
  # adding world map with grey fill and low opacity
  geom_sf(data = countries, fill = "grey", alpha = 0.3) +
  # adding scale for size of points
  new_scale("size") +
  # adding points based on coordinates of centroids with size and color based on count of articles in each country
  geom_point(data = centroids_df, aes(x = x, y = y, size = Count, color = Count), alpha = 0.7) +
  # setting size of points to range from 1 to 10
  scale_size(range = c(1, 10), name = "Count") +
  # setting color of points to range from blue to red
  scale_color_gradient(low = "blue", high = "red", name = "Count") +
  # removing legend for size of points
  theme_void() +
  # setting right projection for world map
  coord_sf() +
  # adding title to plot 
  labs(title = "World Map of Studies in Smart Farming") +
  # adding elements to title
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  # adding legend
  guides(color = guide_legend(override.aes = list(size = 3))) +
  # setting size of legend to false
  guides(size = FALSE)
```

```{r centroid map}
# displaying centroid map
centroid_map
```
